---
title: "Genomic and structural properties of GmPR-1 genes"
author: "Fabrício Almeida-Silva"
date: "3/7/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE,
                      warning=FALSE, 
                      message=FALSE,
                      tidy.opts=list(width.cutoff=60),
                      tidy=TRUE)
```

## Chromosomal distribution


```{r chr_dist}
library(here)
library(tidyverse)
seqs_pr1 <- Biostrings::readAAStringSet(
    here("products", "result_files", "pr1_nostopcodon.fa")
)

genes_and_chrom <- data.frame(Gene=sort(names(seqs_pr1)), 
                              Chr=sort(names(seqs_pr1))) %>%
  mutate(Chr = str_replace(Chr, "Glyma\\.", "")) %>%
  mutate(Chr = str_replace(Chr, "G.*", "")) %>%
  janitor::tabyl(Chr) %>%
  arrange(-n)
genes_and_chrom
```


## GO annotation

Here, we will download GO annotation from PLAZA to see which GO terms are attributed to GmPR-1 genes.

```{r go_annot}
go <- as.data.frame(readr::read_tsv("ftp://ftp.psb.ugent.be/pub/plaza/plaza_public_dicots_04/GO/go.gma.csv.gz", skip=8))
names(go)[1] <- "Gene_ID"
go_pr1 <- go[go$Gene_ID %in% names(seqs_pr1), ]

# Create a data frame with GO terms and their associated genes separated by comma
unique_goterms <- go_pr1 %>%
  select(Gene_ID, go, description) %>%
  unique() %>%
  group_by(go, description) %>%
  summarise(Genes = str_c(Gene_ID, collapse = ", ")) %>%
  mutate(n = str_count(Genes, "Glyma")) %>%
  arrange(-n)

# Manually remove redudant terms (semantic similarity + overlap of associated genes)
print(unique_goterms, n=Inf, width=10)
filtered_goterms <- unique_goterms[c(2:4,7:8,14,16,18:21,23:32,37,
                                     45,48,52,58), ]
filtered_goterms <- filtered_goterms[-c(18,21,24,26), ]

#----Ecport data as a table----
write_tsv(filtered_goterms, file=here("products", "tables", "go_table.tsv"))
```


## Structural analysis

### Homology modelling of PR-1 3D structures

Protein 3D structures were modelled with Phyre2 web server, which performs PSI-BLAST searches against the PDB database and performs homology modelling with MODELLER. 

### Secondary structure analysis

The secondary structure analysis was automatically performed in Phyre2 and exported as a detailed report.

```{r secondary_structure}
second.structure <- data.frame(Genes=c("Glyma.15G062300", "Glyma.17G066100",
                                       "Glyma.15G062900", "Glyma.01G180400",
                                       "Glyma.02G060800", "Glyma.07G186200",
                                       "Glyma.07G186300", "Glyma.07G186400",
                                       "Glyma.10G047000", "Glyma.13G094100",
                                       "Glyma.13G094200", "Glyma.13G251600",
                                       "Glyma.13G251700", "Glyma.13G252000",
                                       "Glyma.13G252300", "Glyma.13G252400",
                                       "Glyma.13G252500", "Glyma.13G252600",
                                       "Glyma.15G062400", "Glyma.15G062500", 
                                       "Glyma.15G062700", "Glyma.15G062800",
                                       "Glyma.16G143300", "Glyma.17G066000"),
                               Alpha_helix=c(40, 43, 35, 26,
                                             35, 38, 39, 36, 
                                             22, 32, 39, 38,
                                             38, 38, 38, 38,
                                             59, 40, 39, 38,
                                             39, 40, 36, 37),
                               Beta_sheet=c(17, 11, 25, 21,
                                            16, 19, 18, 13,
                                            13, 16, 15, 17,
                                            17, 17, 17, 17,
                                            3, 16, 21, 20,
                                            18, 17, 16, 17),
                               Disordered=c(17, 19, 12, 28,
                                            22, 19, 19, 28,
                                            48, 32, 25, 21,
                                            21, 21, 21, 21,
                                            23, 14, 15, 15,
                                            17, 20, 21, 22),
                               TM_helix=c(10, 0, 0, 0,
                                          0, 10, 0, 9,
                                          0, 8, 9, 9,
                                          9, 9, 9, 9,
                                          0, 10, 10, 10,
                                          10, 9, 9, 0)) 
second.structure <- second.structure[order(second.structure$Genes), ]
```

### Physicochemical properties of PR-1

```{r}
library(Peptides)
seqs_pr1 <- seqs_pr1[sort(names(seqs_pr1))]
pi <- pI(as.character(seqs_pr1))
mw <- mw(as.character(seqs_pr1)) / 1000
hydrophibicity <- hydrophobicity(as.character(seqs_pr1))
```

### Subcellular localization and signal peptide

The presence/absence of signal peptides was analyzed with SignalP 5.0 (http://www.cbs.dtu.dk/services/SignalP/) and the prediction summary was downloaded for analysis in R.

```{r}
#----Signal peptide analysis----
signalp <- read.csv(
    here("products", "result_files", "SignalP5.0", "PR1_signalp5.0.txt"), 
    header=TRUE, sep="\t", skip=1
    )

# Create a table containing cleavavge sites of signal peptides
sptable <- signalp
sptable$SP <- sapply(strsplit(sptable$CS.Position, ": "), `[`, 2)
sptable <- sptable[, c(1, 6)]
sptable$SP <- gsub("\\..*", "", sptable$SP)
names(sptable) <- c("Gene_ID", "SP")
sptable <- sptable[order(sptable$Gene_ID), ]
head(sptable)

#----Subcellular localization analysis----
# Load Soybase annotation
load(here("data", "soybase_annotation_tidy.rda"))
cc <- soybase_annotation
cc <- cc %>% 
  filter(Gene %in% names(seqs_pr1)) %>%
  select(Gene, gocc_des) %>%
  distinct() %>%
  arrange(Gene) %>%
  mutate(gocc_des = str_replace(gocc_des, "extracellular region", "Apoplast"))
  as.data.frame()
```


### Genomic properties of PR-1

```{r exon_no_length}
# Stats
exon_no <- unlist(lapply(gene.str.exonlist, nrow))
seqlength <- width(seqs_pr1)
```

Summarizing it all in a table.

```{r table1_paper}
table1 <- data.frame("Gene_ID" = names(seqs_pr1),
                     "Gene_name" = paste0("GmPR-1-", 1:length(seqs_pr1)),
                     "Exon_no" = exon_no,
                     "Protein length (aa)" = seqlength,
                     "pI" = signif(pi, 2),
                     "MW (kDa)" = round(mw),
                     "Hydrophobicity" = signif(hydrophibicity, 2),
                     "SL" = "Apoplast",
                     "SP" = sptable$SP,
                     stringsAsFactors=FALSE)
table1 <- merge(table1, second.structure, by.x="Gene_ID", by.y="Genes")
table1

write.table(
    table1, 
    file=here("products", "tables", "table1.txt"), 
    sep=",", quote=FALSE, row.names=FALSE
    )
```


## Session information

This document was created under the following conditions:

```{r}
sessionInfo()
```